##PeerAssessment 1
###Loading and preprocessing the data

```{r ,echo=TRUE}
setwd("C:/DataSciencecoursera/Reproducible Research/RepData_PeerAssessment1")
data_set<-read.csv("activity.csv",header=TRUE,sep=",")

```
###What is mean total number of steps taken per day?

```{r ,echo=TRUE} 
#calculate the total steps taken each day by applying sum function, group by date
totalSteps_eachDay<-aggregate (data_set$steps,list(date=data_set$date),FUN=sum,na.rm=FALSE)

```
Histogram of the total number of steps taken each day
```{r ,echo=TRUE} 
hist(totalSteps_eachDay[,2],main="Total Steps Taken Each Day",xlab = "Steps",ylab="Frequency")
```




Mean total number of steps taken per day
```{r ,echo=TRUE} 
#calculate the mean steps taken per day
mean(as.numeric(totalSteps_eachDay[,2]),na.rm=TRUE)
```

Median total number of steps taken per day
```{r ,echo=TRUE} 
#calculate the median of  steps taken per day
median(as.numeric(totalSteps_eachDay[,2]),na.rm=TRUE)
```
###What is the average daily activity pattern?
Time series plot of the 5-minute interval and the average number of steps taken, averaged across all days
```{r ,echo=TRUE} 
#Calculate the average (mean) steps for each day by applying mean function, group by interval
meanOf_StepsInInterval<-aggregate (data_set$steps,list(interval=data_set$interval),FUN=mean,na.rm=TRUE)

plot(meanOf_StepsInInterval,xlab="Time Interval",ylab="Average Steps",main="Time Series Plot",type='l')
```



5-minute interval, on average across all the days that contains the maximum number of steps
```{r ,echo=TRUE} 
#get the row number that has max steps
rowID_WithMaxValue<-which.max(meanOf_StepsInInterval$x)
#get the interval with max steps using the row id
meanOf_StepsInInterval[rowID_WithMaxValue,1]
```
###Imputing missing values
Total number of missing values in the dataset 
```{r ,echo=TRUE}
#Calcuate the count of na's in the data set
length(which(is.na(data_set)))
```
 Fill in all of the missing values in the dataset with the mean for that 5-minute interval
```{r ,echo=TRUE}
#find the total rows in main dataset
total_Rows<-nrow(data_set)
#create a new dataframe
imputed_DataSet<-data_set
#Loop over the main dataset
for (i in 1:total_Rows)
{
  # check if there is NA     
  if (is.na(data_set[i,1]) == "TRUE") {    
    #get the mean 
    req_Mean<-meanOf_StepsInInterval[grep(data_set[i,3],meanOf_StepsInInterval[,1],ignore.case=T),]   
    #replace the NA with the interval mean
    imputed_DataSet[i,1]<-req_Mean[1,2]
  } 
}

#Calculate the total # of steps taken each day on the imputed data frame

totalImputedSteps_eachDay<-aggregate (imputed_DataSet$steps,list(date=imputed_DataSet$date),FUN=sum,na.rm=FALSE)
```
Histogram of the total number of steps taken each day with Imputed data
```{r ,echo=TRUE}
#Plot histogram of total Imputed steps each day
hist(totalImputedSteps_eachDay[,2],main="Total Imputed Steps Taken Each Day",xlab = "Steps",ylab="Frequency")
```




Mean of the total number of steps taken each day with Imputed data
```{r ,echo=TRUE}
#calculate the mean of Imputed steps taken per day
mean(as.numeric(totalImputedSteps_eachDay[,2]),na.rm=TRUE)
```
Median of the total number of steps taken each day with Imputed data
```{r ,echo=TRUE}
#calculate the median of  Imputed steps taken per day
median(as.numeric(totalImputedSteps_eachDay[,2]),na.rm=TRUE)
```
After replacing the NA, the mean is the same but the median is different

###Are there differences in activity patterns between weekdays and weekends?
Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r ,echo=TRUE}
#create dataset with new variable stating if the date is weekday or weekend.
imputed_DayDataSet<-imputed_DataSet
imputed_DayDataSet$date <- as.Date(imputed_DayDataSet$date, "%Y-%m-%d")
days<-weekdays(imputed_DayDataSet$date)
daylevel <- vector()
for (i in 1:nrow(imputed_DayDataSet)) {
  if (days[i] == "Saturday") {
    daylevel[i] <- "Weekend"
  } else if (days[i] == "Sunday") {
    daylevel[i] <- "Weekend"
  } else {
    daylevel[i] <- "Weekday"
  }
}

imputed_DayDataSet$daylevel <- daylevel
imputed_DayDataSet$daylevel <- factor(imputed_DayDataSet$daylevel)
table(imputed_DayDataSet$daylevel)
```
Panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r ,echo=TRUE}
library(lattice)
stepsByDay <- aggregate(steps ~ interval + daylevel, data = imputed_DayDataSet, mean)
xyplot(steps ~ interval | daylevel, stepsByDay, type = "l", layout = c(1, 2), 
       xlab = "Interval", ylab = "Number of steps")
```
